---
title: "Fish Density"
author: "Kaitlin Rommelfanger"
date: "2023-09-24"
output: html_document
editor_options: 
  chunk_output_type: console
---
3) Seagrass NMDS
  Permanova
  Simper
4) Fish Density
5) T test on fish density
6) Clean up Plots
  Add Stress
  Change up Colors for Depth and % Halophila
7) plot top seven species listed by depth, and run a t test at the two separate depths.  
  
  
  
```{r Fish Density}
library(tidyverse)
#we have the number of fish per transect in we fixed it analyze me, we need to calculate density
fish<-read.csv("Data/WeFixedItAnalyzeMe.csv")
View(fish)

#our transects were 100m2 so we divided the total number of fish by 100, to get fish density per transect
fish<-fish%>%mutate("density"=(N/100))

#now we need to add a dcat column, dividing depth into <15 and >15 m

fish$dcat<-case_when(fish$depth<15 ~ "1", fish$depth>15 ~"2")

# now I gotta remember how to run a t test...

t.test(density~dcat, data=fish)

#	Welch Two Sample t-test

#data:  density by dcat
#t = -1.4718, df = 52.253, p-value = 0.1471
#alternative hypothesis: true difference in means between group 1 and group 2 is not equal to 0
#95 percent confidence interval:
# -0.18131652  0.02787114
#sample estimates:
#mean in group 1 mean in group 2 
#      0.1997059       0.2764286 



#now looking at fish density across different percentages of halophila
grass<-read.csv("Data/WeFixedItAnalyzeMe.csv")

grass<-grass%>%mutate(totalgrass=(PCovHalophila+PCovThalassia+PCovSyringodium+PCovHalodule))%>%
  mutate(PerHalTotal=(PCovHalophila/totalgrass))

#turn nas from dividing by 0 to zeros 
grass[is.na(grass)]<-0

#make two categories for when halophila is the only seagrass and when there are other seagrasses present 
grass$halcat <- case_when(
  grass$PerHalTotal<1 ~ "1",
  grass$PerHalTotal==1 ~ "2")

#make density column
grass<-grass%>%mutate("density"=(N/100))

#now we run our t test 
t.test(density~halcat, data=grass)

#	Welch Two Sample t-test

#data:  density by halcat
#t = -2.5139, df = 57.795, p-value = 0.01475
#alternative hypothesis: true difference in means between group 1 and group 2 is not equal to 0
#95 percent confidence interval:
# -0.20114487 -0.02280751
#sample estimates:
#mean in group 1 mean in group 2 
 #     0.1585000       0.2704762 

```

```{r Stefs Code, slightly tweaked}
##load libraries##
library(vegan)
library(DHARMa)
library(tidyverse)
library(dplyr)

pcov<-read.csv("Data/WeFixedItAnalyzeMe.csv")

#Add new column for % halophila/ % total seagrass cover#
pcov<-pcov %>% 
  mutate(totalgrass= (pcov$PCovHalophila+pcov$PCovThalassia+pcov$PCovHalodule+pcov$PCovSyringodium)) %>% 
  mutate(PerHalTotal=(pcov$PCovHalophila/totalgrass))

##Making the NAs 0
pcov[is.na(pcov)] <- 0

#rename the dataframe
depth_hal<-pcov

##Make the depth bin categories##
depth_hal<- depth_hal%>% 
       mutate(dhcat = case_when((depth >= 15.0) ~ 1,
                                (depth <= 15.0 & PerHalTotal == 1.0) ~ 2,
                                (depth <= 15.0 & PerHalTotal < 1.0) ~ 3 ))


##Read in the fish data##
AllMerged <- read_csv("AllMerged.csv")

##making a dataframe
AllMerged$Family<-as.factor(AllMerged$Family)

#determine the total number of fish of each family that are on each transect to make dataframe for nMDS
dhnMDSframe<-AllMerged %>%
  group_by(Transect)%>%
  count(Family)
 
#we want wide instead of long
dhnMDSframe <- dhnMDSframe %>%
  pivot_wider(names_from = Family , values_from = n)

##Turning NAs to 0s
dhnMDSframe[is.na(dhnMDSframe)] <- 0

##Merging the two dataframes to make an nMDS dataframe, that contains 
dhnMDSframe = merge(dhnMDSframe, depth_hal, by = "Transect")

##Removing unnecessary columns
#dhnMDSframe <- dhnMDSframe[,-21:-37]
#dhnMDSframe <- dhnMDSframe[,-1]

#Select for only the columns that contain the families so we can run the nMDS
HnMDSframe<-dhnMDSframe%>%select(Aulostomidae:Haemulidae)

#make a frame that has transect and our cat column in it to use for plotting purposes  
plotframe<-dhnMDSframe%>%select(Aulostomidae:Haemulidae, Transect,dhcat)

#run our nMDS
family_NMDS <- metaMDS(HnMDSframe, k = 2, trymax = 100) 
family_NMDS
#Stress:     0.2339253 




trans.cont.dh <- as.data.frame(family_NMDS$points)
trans.cont.dh$dhcat <- dhnMDSframe[,20]

 

##Making new ellipses for Halophila
DHord <- ordiellipse(family_NMDS, trans.cont.dh$dhcat, conf = 0.95, label = TRUE)

 

##Calling the hidden vegan function##
veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
{
  theta <- (0:npoints) * 2 * pi/npoints
  Circle <- cbind(cos(theta), sin(theta))
  t(center + scale * t(Circle %*% chol(cov)))
}

 

 

trans.cont.dh$dhcat <- factor(trans.cont.dh$dhcat)

 

 

 

##Making a new dataframe for the Halophila ellipses
DH_df_ell <- data.frame()
for(g in levels(trans.cont.dh$dhcat)){
  DH_df_ell <- rbind(DH_df_ell, cbind(as.data.frame(with(trans.cont.dh[trans.cont.dh$dhcat==g,],
                                                               veganCovEllipse(DHord[[g]]$cov,DHord[[g]]$center,DHord[[g]]$scale)))
                                            ,zone=g))
}

 

##Plot
ggplot(trans.cont.dh) +
  geom_point(aes(x = MDS1, y = MDS2, color = as.factor(dhcat)), size = 3) + geom_path(data = DH_df_ell, aes(x = NMDS1, y = NMDS2, group=zone, colour=zone), linetype=1, linewidth=2)+
  theme_classic() +
  labs(color = "cat", fill = "cat") +
  guides(color = guide_legend(nrow = 1), fill = guide_legend(nrow = 1)) +
  theme(panel.border = element_rect(fill = NA),
        legend.position = "bottom")+ geom_text(data = fish_family, aes(x = MDS1, y = MDS2, label = sp), alpha = 0.3)+scale_fill_manual(name = "Depth and Halophila Categories", labels = c("Greater than 15 m ", "Less than 15m and 100% Halophila", "Less than 15m and Less than 100% Halophila"),values = c("#3B9AB2","#EBCC2A", "#F21A00"))+scale_color_manual(name = "Depth", labels = c("Greater than 15 m", "Less than 15m and 100% Halophila", "Less than 15m and Less than 100% Halophila"), values = c("#3B9AB2","#EBCC2A","#F21A00"))

 

DH_dissim_bray<- dhnMDSframe
dh_bray <- vegdist(DH_dissim_bray[,1:19], method = "bray")
    dh_bray

 

    ##running the depth 1-Way PERMANOVA##
    dh_Strata <- adonis2(dh_bray ~ DH_dissim_bray$dhcat, data = DH_dissim_bray[,1:19])
    dh_Strata

    ## DH simper
    dh_simper <- simper(DH_dissim_bray[,1:19], permutations = 999, group = (DH_dissim_bray$dhcat))
  summary(dh_simper)

    ##Results
  #Contrast: 3_1

 

           #     average      sd   ratio     ava     avb cumsum     p    
#Labridae        0.18416 0.18313 1.00570 8.42100 1.28600  0.232 0.426    
#Lutjanidae      0.16823 0.17510 0.96080 0.05300 8.67900  0.445 0.001 ***
#Gobiidae        0.12732 0.13045 0.97600 1.63200 5.71400  0.605 0.630    
#Serranidae      0.10689 0.09100 1.17470 1.63200 4.75000  0.740 0.016 *  
#Scaridae        0.07222 0.10119 0.71370 2.21100 1.60700  0.831 0.736    
#Labrisomidae    0.03837 0.05042 0.76100 0.21100 1.39300  0.880 0.044 *  
#Carangidae      0.02938 0.11892 0.24700 0.00000 2.50000  0.917 0.338    
#Monacanthidae   0.02007 0.02516 0.79740 0.26300 0.67900  0.942 0.796    
#Tetraodontidae  0.01758 0.03660 0.48040 0.47400 0.17900  0.965 0.139    
#Pomacentridae   0.01066 0.03055 0.34890 0.52600 0.10700  0.978 0.915    
#Clinidae        0.00684 0.03665 0.18650 0.00000 0.25000  0.987 0.313    
#Haemulidae      0.00414 0.01542 0.26820 0.05300 0.14300  0.992 0.260    
#Aulostomidae    0.00229 0.00801 0.28580 0.05300 0.03600  0.995 0.223    
#Opistognathidae 0.00215 0.01238 0.17390 0.00000 0.03600  0.997 0.303    
#Acanthuridae    0.00134 0.00709 0.18910 0.00000 0.07100  0.999 0.945    
#Dactylopteridae 0.00074 0.00394 0.18850 0.00000 0.03600  1.000 0.306    
#Holocentridae   0.00000 0.00000     NaN 0.00000 0.00000  1.000 0.745    
#Syngnathidae    0.00000 0.00000     NaN 0.00000 0.00000  1.000 0.938    
#Chaetodontidae  0.00000 0.00000     NaN 0.00000 0.00000  1.000 0.773    

 

  # Fish in the family Lutjanidae (aka snappers) account for 21.3% of the   dissimilarity between the greater than 15m category and the less than 15 m with less than 100% Halophila category (p=.001)

 

#Contrast: 3_2

 

#                 average       sd    ratio      ava      avb cumsum     p    
#Labridae       0.26277  0.20726  1.26780  8.42100 11.20000  0.376 0.001 ***
#Gobiidae         0.14299  0.13176  1.08520  1.63200  5.40000  0.581 0.216  
#Scaridae         0.09976  0.10690  0.93310  2.21100  3.53300  0.724 0.059 .
#Serranidae       0.06324  0.07484  0.84490  1.63200  1.53300  0.814 0.997  #Pomacentridae    0.03127  0.06064  0.51570  0.52600  1.20000  0.859 0.008 ** #Monacanthidae    0.02600  0.03429  0.75850  0.26300  0.80000  0.896 0.071 #Tetraodontidae   0.01744  0.04027  0.43310  0.47400  0.13300  0.921 0.286   #Labrisomidae     0.01592  0.02906  0.54790  0.21100  0.33300  0.944 0.994  #Lutjanidae       0.01403  0.02363  0.59370  0.05300  0.46700  0.964 1.000  #Syngnathidae    0.00767  0.02429  0.31570  0.00000  0.20000  0.975 0.001 ***
#Acanthuridae     0.00674  0.01802  0.37390  0.00000  0.46700  0.985 0.119  #Chaetodontidae   0.00473  0.02025  0.23350  0.00000  0.06700  0.992 0.007 ** #Holocentridae    0.00270  0.01072  0.25200  0.00000  0.06700  0.995 0.006 ** #Haemulidae       0.00165  0.00772  0.21390  0.05300  0.00000  0.998 0.755  #Aulostomidae     0.00154  0.00712  0.21590  0.05300  0.00000  1.000 0.581  #Clinidae         0.00000  0.00000      NaN  0.00000  0.00000  1.000 0.546  #Dactylopteridae  0.00000  0.00000      NaN  0.00000  0.00000  1.000 0.560  #Carangidae       0.00000  0.00000      NaN  0.00000  0.00000  1.000 0.962  #Opistognathidae  0.00000  0.00000      NaN  0.00000  0.00000  1.000 0.532    

 

#Contrast: 1_2

 

#                 average       sd    ratio      ava      avb cumsum     p  
#Labridae         0.17351  0.16450  1.05480  1.28600 11.20000  0.225 0.608  
#Lutjanidae       0.14687  0.15930  0.92190  8.67900  0.46700  0.416 0.106  
#Gobiidae         0.13724  0.12792  1.07280  5.71400  5.40000  0.594 0.304  
#Serranidae       0.09477  0.08432  1.12390  4.75000  1.53300  0.717 0.353  
#Scaridae         0.07763  0.09470  0.81970  1.60700  3.53300  0.817 0.485  
#Labrisomidae     0.03307  0.04489  0.73680  1.39300  0.33300  0.860 0.390  
#Carangidae       0.02686  0.11117  0.24160  2.50000  0.00000  0.895 0.465  
#Monacanthidae    0.02318  0.02741  0.84550  0.67900  0.80000  0.925 0.303  
#Pomacentridae    0.02099  0.04699  0.44670  0.10700  1.20000  0.952 0.249  
#Tetraodontidae   0.00721  0.01627  0.44290  0.17900  0.13300  0.962 0.964  
#Acanthuridae     0.00677  0.01657  0.40860  0.07100  0.46700  0.971 0.067 .
#Clinidae         0.00591  0.03271  0.18070  0.25000  0.00000  0.978 0.504  
#Syngnathidae     0.00590  0.01911  0.30870  0.00000  0.20000  0.986 0.089 .
#Chaetodontidae   0.00322  0.01443  0.22290  0.00000  0.06700  0.990 0.229  
#Haemulidae       0.00239  0.01288  0.18530  0.14300  0.00000  0.993 0.641  
#Holocentridae    0.00207  0.00852  0.24340  0.00000  0.06700  0.996 0.214  
#Opistognathidae  0.00175  0.01085  0.16150  0.03600  0.00000  0.998 0.470  
#Aulostomidae     0.00082  0.00454  0.18110  0.03600  0.00000  0.999 0.859  
#Dactylopteridae  0.00066  0.00356  0.18430  0.03600  0.00000  1.000 0.522  


```

```{r final nMDS}
library(tidyverse)
library(vegan)

pcov<-read.csv("Data/WeFixedItAnalyzeMe.csv")

#Add new column for % halophila/ % total seagrass cover#
pcov<-pcov %>% 
  mutate(totalgrass= (pcov$PCovHalophila+pcov$PCovThalassia+pcov$PCovHalodule+pcov$PCovSyringodium)) %>% 
  mutate(PerHalTotal=(pcov$PCovHalophila/totalgrass))

##Making the NAs 0
pcov[is.na(pcov)] <- 0

#rename the dataframe
depth_hal<-pcov

##Make the depth bin categories##
depth_hal<- depth_hal%>% 
       mutate(dhcat = case_when((depth >= 15.0) ~ 1,
                                (depth <= 15.0 & PerHalTotal == 1.0) ~ 2,
                                (depth <= 15.0 & PerHalTotal < 1.0) ~ 3 ))


##Read in the fish data##
AllMerged <- read.csv("AllMerged.csv")

##making a dataframe
AllMerged$Family<-as.factor(AllMerged$Family)

#determine the total number of fish of each family that are on each transect to make dataframe for nMDS
dhnMDSframe<-AllMerged %>%
  group_by(Transect)%>%
  count(Family)
 
#we want wide instead of long
dhnMDSframe <- dhnMDSframe %>%
  pivot_wider(names_from = Family , values_from = n)

##Turning NAs to 0s
dhnMDSframe[is.na(dhnMDSframe)] <- 0

##Merging the two dataframes to make an nMDS dataframe, tabular by transect 
dhnMDSframe = merge(dhnMDSframe, depth_hal, by = "Transect")

##Removing unnecessary columns
#dhnMDSframe <- dhnMDSframe[,-21:-37]
#dhnMDSframe <- dhnMDSframe[,-1]

#Select for only the columns that contain the families so we can run the nMDS
  HnMDSframe<-dhnMDSframe%>%
    select(`Aulostomidae `:Haemulidae)

#make a frame that has transect and our cat column in it to use for plotting purposes  
plotframe<-dhnMDSframe%>%
  select(`Aulostomidae `:Haemulidae, Transect,dhcat)

#now run the nMDS 
final_NMDS <- metaMDS(HnMDSframe, k = 2, trymax = 100) 
final_NMDS
#Stress:     0.2339253 

stressplot(final_NMDS)
#non metric r squared 0.945
#linear fit r2=0.72

##pulling the species locations out of the NMDS (x and y)
fam <- as.data.frame(final_NMDS$species)

# this line is pulling the species names and linking them to thier location
fam$sp <- rownames(final_NMDS$species)


#this is pulling the points for each individual transect out of the nmds
  transect <- as.data.frame(final_NMDS$points)

#we are pulling the dhcat out of our plotframe that we made earlier so we can make the polygons depth/percent halophila
  transect$dhcat <- plotframe[,21] 

#need to keep this code because you cant pull the points in the next line unless if this ordiplot exists
ordiplot(final_NMDS, type = "none") 

#this is storing the info to make the 95% confidence intervals as an object 
finallord<-ordiellipse(final_NMDS, transect$dhcat, conf = 0.95, label = TRUE)

#hidden vegan function that calculates ellipses
veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
{
  theta <- (0:npoints) * 2 * pi/npoints
  Circle <- cbind(cos(theta), sin(theta))
  t(center + scale * t(Circle %*% chol(cov)))
}

#convert zone to a factor
transect$dhcat <- factor(transect$dhcat)

#create a dataframe of points that will draw the ellipses
final_df_ell <- data.frame()

for(g in levels(transect$dhcat)){
  final_df_ell <- rbind(final_df_ell, cbind(as.data.frame(with(transect[transect$dhcat==g,],
                                                   veganCovEllipse(finallord[[g]]$cov,finallord[[g]]$center,finallord[[g]]$scale)))
                                ,zone=g))
}

#plot
ggplot(transect) +
  geom_point(aes(x = MDS1, y = MDS2, color = as.factor(dhcat)), size = 3) + geom_path(data = final_df_ell, aes(x = NMDS1, y = NMDS2, group = zone))+
  theme_classic() +
  labs(color = "dhcat", fill = "dhcat") +
  guides(color = guide_legend(nrow = 1), fill = guide_legend(nrow = 1)) +
theme(panel.border = element_rect(fill = NA),
legend.position = "bottom")+ geom_text(data = fam, aes(x = MDS1, y = MDS2, label = sp), alpha = 0.3)+scale_fill_manual(name = "Category", labels = c("Greater than 15m", "100% Halophila less than 15m","Less than 100% Halophila and Less than 15m"),values = c("#3B9AB2","#EBCC2A","purple"))+scale_color_manual(name = "Category", labels = c("Greater than 15m", "100% Halophila less than 15m","Less than 100% Halophila and Less than 15m"), values = c("#3B9AB2","#EBCC2A","purple"))


#now run post hoc stuff to see if we have significant differences between the groups

final_dissim_bray<- HnMDSframe


final_bray <- vegdist(final_dissim_bray, method = "bray")
    final_bray

 

    ##running the depth 1-Way PERMANOVA##
    dh_Strata <- adonis2(final_bray ~ plotframe$dhcat, data = final_dissim_bray)
    dh_Strata
    
    
#there is a significant difference 
    
     ## DH simper
       dh_simper <- simper(final_dissim_bray, permutations = 999, group = (transect$dhcat))
  summary(dh_simper)
  
  
#One= greater than 15m
#Two= Less than 15m and all seagrass present is Halophila
#Three= Less than 15m, not all seagrass is Halophila
  
  
  
#3_1 is significant, simper shows that shallow seagrass containing native species has significantly different Lutjanidae, Serranidae, Labrisomidae (? stef and aves got significant I did not) than deep
  
#3_2 is significant simper shows that seagrass with native is significantly different Labridae, Pomacentridae, Syngnathidae, Chaetodontidae, Holocentridae than shallow 100% halophila
  
  
#1_2 no significant families different in shallow and deep halophila
  
  ((depth >= 15.0) ~ 1,
                                (depth <= 15.0 & PerHalTotal == 1.0) ~ 2,
                                (depth <= 15.0 & PerHalTotal < 1.0) ~ 3 ))


Contrast: 3_1 

                average      sd   ratio     ava     avb cumsum     p   
Labridae        0.18416 0.18313 1.00570 8.42100 1.28600  0.232 0.424   
Lutjanidae      0.16823 0.17510 0.96080 0.05300 8.67900  0.445 0.002 **
Gobiidae        0.12732 0.13045 0.97600 1.63200 5.71400  0.605 0.673   
Serranidae      0.10689 0.09100 1.17470 1.63200 4.75000  0.740 0.011 * 
Scaridae        0.07222 0.10119 0.71370 2.21100 1.60700  0.831 0.769   
Labrisomidae    0.03837 0.05042 0.76100 0.21100 1.39300  0.880 0.056 . 
Carangidae      0.02938 0.11892 0.24700 0.00000 2.50000  0.917 0.323   
Monacanthidae   0.02007 0.02516 0.79740 0.26300 0.67900  0.942 0.803   
Tetraodontidae  0.01758 0.03660 0.48040 0.47400 0.17900  0.965 0.118   
Pomacentridae   0.01066 0.03055 0.34890 0.52600 0.10700  0.978 0.913   
Clinidae        0.00684 0.03665 0.18650 0.00000 0.25000  0.987 0.329   
Haemulidae      0.00414 0.01542 0.26820 0.05300 0.14300  0.992 0.233   
Aulostomidae    0.00229 0.00801 0.28580 0.05300 0.03600  0.995 0.222   
Opistognathidae 0.00215 0.01238 0.17390 0.00000 0.03600  0.997 0.312   
Acanthuridae    0.00134 0.00709 0.18910 0.00000 0.07100  0.999 0.833   
Dactylopteridae 0.00074 0.00394 0.18850 0.00000 0.03600  1.000 0.312   
Holocentridae   0.00000 0.00000     NaN 0.00000 0.00000  1.000 0.741   
Syngnathidae    0.00000 0.00000     NaN 0.00000 0.00000  1.000 0.931   
Chaetodontidae  0.00000 0.00000     NaN 0.00000 0.00000  1.000 0.760   
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Contrast: 3_2 

                 average       sd    ratio      ava      avb cumsum
Labridae         0.26329  0.20737  1.26970  8.42100 11.20000  0.377
Gobiidae         0.14315  0.13177  1.08640  1.63200  5.40000  0.583
Scaridae         0.09991  0.10701  0.93360  2.21100  3.53300  0.726
Serranidae       0.06338  0.07501  0.84500  1.63200  1.53300  0.817
Pomacentridae    0.03189  0.06281  0.50780  0.52600  1.20000  0.862
Monacanthidae    0.02603  0.03431  0.75880  0.26300  0.80000  0.900
Tetraodontidae   0.01749  0.04037  0.43320  0.47400  0.13300  0.925
Labrisomidae     0.01595  0.02911  0.54790  0.21100  0.33300  0.948
Lutjanidae       0.01403  0.02364  0.59380  0.05300  0.46700  0.968
Syngnathidae     0.00767  0.02429  0.31570  0.00000  0.20000  0.979
Chaetodontidae   0.00473  0.02025  0.23350  0.00000  0.06700  0.986
Acanthuridae     0.00403  0.01521  0.26510  0.00000  0.40000  0.991
Holocentridae    0.00283  0.01128  0.25100  0.00000  0.06700  0.995
Haemulidae       0.00166  0.00774  0.21390  0.05300  0.00000  0.998
Aulostomidae     0.00154  0.00713  0.21590  0.05300  0.00000  1.000
Clinidae         0.00000  0.00000      NaN  0.00000  0.00000  1.000
Dactylopteridae  0.00000  0.00000      NaN  0.00000  0.00000  1.000
Carangidae       0.00000  0.00000      NaN  0.00000  0.00000  1.000
Opistognathidae  0.00000  0.00000      NaN  0.00000  0.00000  1.000
                    p    
Labridae        0.001 ***
Gobiidae        0.234    
Scaridae        0.064 .  
Serranidae      0.999    
Pomacentridae   0.006 ** 
Monacanthidae   0.083 .  
Tetraodontidae  0.290    
Labrisomidae    0.998    
Lutjanidae      1.000    
Syngnathidae    0.002 ** 
Chaetodontidae  0.006 ** 
Acanthuridae    0.310    
Holocentridae   0.011 *  
Haemulidae      0.765    
Aulostomidae    0.533    
Clinidae        0.584    
Dactylopteridae 0.528    
Carangidae      0.968    
Opistognathidae 0.557    
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Contrast: 1_2 

                 average       sd    ratio      ava      avb cumsum
Labridae         0.17391  0.16461  1.05650  1.28600 11.20000  0.226
Lutjanidae       0.14713  0.15956  0.92210  8.67900  0.46700  0.417
Gobiidae         0.13749  0.12816  1.07280  5.71400  5.40000  0.595
Serranidae       0.09498  0.08452  1.12380  4.75000  1.53300  0.718
Scaridae         0.07773  0.09483  0.81960  1.60700  3.53300  0.819
Labrisomidae     0.03315  0.04500  0.73670  1.39300  0.33300  0.862
Carangidae       0.02689  0.11126  0.24170  2.50000  0.00000  0.897
Monacanthidae    0.02321  0.02744  0.84580  0.67900  0.80000  0.927
Pomacentridae    0.02139  0.04861  0.44000  0.10700  1.20000  0.955
Tetraodontidae   0.00722  0.01630  0.44280  0.17900  0.13300  0.964
Clinidae         0.00592  0.03278  0.18070  0.25000  0.00000  0.972
Syngnathidae     0.00590  0.01911  0.30870  0.00000  0.20000  0.980
Acanthuridae     0.00479  0.01498  0.31980  0.07100  0.40000  0.986
Chaetodontidae   0.00322  0.01443  0.22290  0.00000  0.06700  0.990
Haemulidae       0.00239  0.01290  0.18530  0.14300  0.00000  0.993
Holocentridae    0.00215  0.00889  0.24210  0.00000  0.06700  0.996
Opistognathidae  0.00176  0.01088  0.16170  0.03600  0.00000  0.998
Aulostomidae     0.00082  0.00455  0.18110  0.03600  0.00000  0.999
Dactylopteridae  0.00066  0.00357  0.18430  0.03600  0.00000  1.000
                    p  
Labridae        0.574  
Lutjanidae      0.105  
Gobiidae        0.327  
Serranidae      0.353  
Scaridae        0.539  
Labrisomidae    0.412  
Carangidae      0.469  
Monacanthidae   0.284  
Pomacentridae   0.244  
Tetraodontidae  0.966  
Clinidae        0.536  
Syngnathidae    0.086 .
Acanthuridae    0.166  
Chaetodontidae  0.233  
Haemulidae      0.657  
Holocentridae   0.218  
Opistognathidae 0.471  
Aulostomidae    0.855  
Dactylopteridae 0.536  
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
Permutation: free
Number of permutations: 999



```

Plot Densities of top 90% families differences
```{r}
#3_1 is significant, simper shows that shallow seagrass containing native species has significantly different Lutjanidae, Serranidae, Labrisomidae (? stef and aves got significant I did not) than deep

#                average      sd   ratio     ava     avb cumsum     p   
#Labridae        0.18416 0.18313 1.00570 8.42100 1.28600  0.232 0.424   
#Lutjanidae      0.16823 0.17510 0.96080 0.05300 8.67900  0.445 0.002 **
#Gobiidae        0.12732 0.13045 0.97600 1.63200 5.71400  0.605 0.673   
#Serranidae      0.10689 0.09100 1.17470 1.63200 4.75000  0.740 0.011 * 
#Scaridae        0.07222 0.10119 0.71370 2.21100 1.60700  0.831 0.769   
#Labrisomidae    0.03837 0.05042 0.76100 0.21100 1.39300  0.880 0.056 . 
#Carangidae      0.02938 0.11892 0.24700 0.00000 2.50000  0.917 0.323 


#plotting densities of top 90% fish shallow not all halophila vs >15m


#we have the number of fish per transect in we fixed it analyze me, we need to calculate density
deepvsnative<-read.csv("Data/WeFixedItAnalyzeMe.csv")


#our transects were 100m2 so we divided the total number of fish by 100, to get fish density per transect
deepvsnative<-deepvsnative%>%mutate("density"=(N/100))

#adding a totalgrass column, sum of all grass present, and calculating the percent of present seagrass that is halophila 
deepvsnative<-deepvsnative %>% 
  mutate(totalgrass= (pcov$PCovHalophila+pcov$PCovThalassia+pcov$PCovHalodule+pcov$PCovSyringodium)) %>% 
  mutate(PerHalTotal=(pcov$PCovHalophila/totalgrass))

#cannot divide by 0, when total cover is zero got nas, change to zeros 
deepvsnative[is.na(deepvsnative)]<-0

##Make the depth bin categories##
deepvsnative<- deepvsnative%>% 
       mutate(dhcat = case_when((depth >= 15.0) ~ 1,
                                (depth <= 15.0 & PerHalTotal == 1.0) ~ 2,
                                (depth <= 15.0 & PerHalTotal < 1.0) ~ 3 ))



#import dataframe
GNativeVsNon<-read.csv("AllMerged.csv")

#count of number of each family on each transect, want tabular by transect, replaced NA with zeros, then selected the families contributing to 90% of the difference

FNativeVsNon<-FDeepVsNative%>%group_by(Transect)%>%
  count(Family)%>%pivot_wider(names_from = Family, values_from = n)%>%
  mutate_all(~replace(., is.na(.), 0))


  FNativeVsNon<-FNativeVsNon%>%select(Transect,Labridae, Lutjanidae, Gobiidae, Serranidae, Scaridae, Labrisomidae, Carangidae,Pomacentridae, Monacanthidae )%>%mutate("Lab"=Labridae/100, 
                                  "Lut"=Lutjanidae/100,
                                  "Gob"=Gobiidae/100,
                                  "Ser"=Serranidae/100,
                                  "Scar"=Scaridae/100,
                                  "Labm"=Labrisomidae/100,
                                  "Car"=Carangidae/100,
                                  "Pom"=Pomacentridae/100,
                                  "Mon"=Monacanthidae/100)
  


#now merge the two dataframes we just want the dhcat from the first dataframe 
FNativeVsNon<-merge(FNativeVsNon, deepvsnative[,c("Transect", "dhcat")], by="Transect")

#select columns I want 
plotframe <- FNativeVsNon%>%select(Transect, Lab, Lut, Gob, Ser, Scar, Labm, Car, Pom, Mon, dhcat)

plotframe <- plotframe%>%
  pivot_longer(cols = Lab:Mon,
               names_to = "Family",
               values_to = "density")%>%
  mutate(dhcat = as.factor(dhcat))

FNativeVsNonPlot <- ggplot(data = plotframe, aes(x = Family, y = density, fill = dhcat))+
  geom_col(position = position_dodge(0.9), width = 0.75)+
  labs(x = "Family", y = "Densities", fill = "Categories")  

FNativeVsNonPlot

```

These are the families contributing to the top 90 percent of difference between groups

Labridae        0.18416 0.18313 1.00570 8.42100 1.28600  0.232 0.424   
Lutjanidae      0.16823 0.17510 0.96080 0.05300 8.67900  0.445 0.002 **
Gobiidae        0.12732 0.13045 0.97600 1.63200 5.71400  0.605 0.673   
Serranidae      0.10689 0.09100 1.17470 1.63200 4.75000  0.740 0.011 * 
Scaridae        0.07222 0.10119 0.71370 2.21100 1.60700  0.831 0.769   
Labrisomidae    0.03837 0.05042 0.76100 0.21100 1.39300  0.880 0.056 . 
Carangidae      0.02938 0.11892 0.24700 0.00000 2.50000  0.917 0.323  

                 average       sd    ratio      ava      avb cumsum
Labridae         0.26329  0.20737  1.26970  8.42100 11.20000  0.377
Gobiidae         0.14315  0.13177  1.08640  1.63200  5.40000  0.583
Scaridae         0.09991  0.10701  0.93360  2.21100  3.53300  0.726
Serranidae       0.06338  0.07501  0.84500  1.63200  1.53300  0.817
Pomacentridae    0.03189  0.06281  0.50780  0.52600  1.20000  0.862
Monacanthidae    0.02603  0.03431  0.75880  0.26300  0.80000  0.900

                 average       sd    ratio      ava      avb cumsum
Labridae         0.17391  0.16461  1.05650  1.28600 11.20000  0.226
Lutjanidae       0.14713  0.15956  0.92210  8.67900  0.46700  0.417
Gobiidae         0.13749  0.12816  1.07280  5.71400  5.40000  0.595
Serranidae       0.09498  0.08452  1.12380  4.75000  1.53300  0.718
Scaridae         0.07773  0.09483  0.81960  1.60700  3.53300  0.819
Labrisomidae     0.03315  0.04500  0.73670  1.39300  0.33300  0.862
Carangidae       0.02689  0.11126  0.24170  2.50000  0.00000  0.897
Monacanthidae    0.02321  0.02744  0.84580  0.67900  0.80000  0.927


```{r Lab KW}
library(FSA)#for DunnPost Hoc
library(rcompanion)#for cldList function
library(tidyverse)

View(plotframe)

#filter for Labridae only 
Lab<-plotframe%>%filter(Family=="Lab")

Lab%>%ggplot(aes(x=Transect, y=density))+
  geom_point()

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Lab)
#p-value = 6.948e-16

hist(Lab$density)

#transform log x+1
Lab$logxplus1<-log(Lab$density+1)

hist(Lab$logxplus1)


#try boxcox transformation
#https://stackoverflow.com/questions/33999512/how-to-use-the-box-cox-power-transformation-in-r

density<-c(Lab[,'density'])
unlist(density)
dhcat<-as.vector(Lab[,'dhcat'])
boxcox(density~dhcat)


# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Lab)

#	Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 17.592, df = 2, p-value =
#0.0001513


#now we run post hoc analysis
LabDunn=dunnTest(density ~ dhcat,
  data = Lab)

LabDunn

#Dunn (1964) Kruskal-Wallis multiple comparison
#  p-values adjusted with the Holm method.

#  Comparison          Z      P.unadj       P.adj
#1      1 - 2 -3.3737694 0.0007414646 0.001482929
#2      1 - 3 -3.5765002 0.0003482251 0.001044675
#3      2 - 3  0.0476598 0.9619873716 0.961987372

Lab1 = LabDunn$res

cldList(comparison = Lab1$Comparison,
        p.value    = Lab1$P.adj,
        threshold  = 0.05)

#  Group Letter MonoLetter
#1     1      a         a 
#2     2      b          b
#3     3      b          b

```

```{r Gob KW}
View(plotframe)

#filter for Labridae only 
Gob<-plotframe%>%filter(Family=="Gob")

hist(Gob)
hist(Gob$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Gob)
#p-value = 5.067e-05


#transform log x+1
Gob$logxplus1<-log(Gob$density+1)

hist(Gob$logxplus1)

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Gob)

#	Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 8.5397, df = 2, p-value = 0.01398


#now we run post hoc analysis
GobDunn=dunnTest(density ~ dhcat,
  data = Gob)

GobDunn

#  Comparison         Z    P.unadj      P.adj
#1      1 - 2 0.1243478 0.90103987 0.90103987
#2      1 - 3 2.7526420 0.00591165 0.01773495
#3      2 - 3 2.2535917 0.02422187 0.04844373

Gob1 = GobDunn$res

cldList(comparison = Gob1$Comparison,
        p.value    = Gob1$P.adj,
        threshold  = 0.05)

#  Group Letter MonoLetter
#1     1      a         a 
#2     2      a         a 
#3     3      b          b
```

```{r Scar KW}
View(plotframe)

#filter for Labridae only 
Scar<-plotframe%>%filter(Family=="Scar")


hist(Scar$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Scar)
#p-value = 0.1001


ScarANOVA <- aov(density ~ dhcat, data = Scar)

# then we can easily pull them out of the output
aov_residuals <- residuals(ScarANOVA)

# let's first look to see if they look approximately normally distributed
hist(aov_residuals)

# then we can test if the residuals are approximately normally distributed with the Shapiro-Wilk Test
shapiro.test(aov_residuals) 
#nope, sad day, nan parametric it is

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Scar)

#	Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 2.0336, df = 2, p-value = 0.3618

```

```{r Ser KW}
View(plotframe)

#filter for Labridae only 
Ser<-plotframe%>%filter(Family=="Ser")


hist(Ser$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Ser)
#p-value = 0.0008415

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Ser)

#	Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 11.466, df = 2, p-value =
#0.003237

#now we run post hoc analysis
SerDunn=dunnTest(density ~ dhcat,
  data = Ser)

SerDunn

#  Comparison          Z     P.unadj       P.adj
#1      1 - 2  2.9840092 0.002844982 0.008534946
#2      1 - 3  2.6040913 0.009211821 0.018423641
#3      2 - 3 -0.5233973 0.600697770 0.600697770

Ser1 = SerDunn$res

cldList(comparison = Ser1$Comparison,
        p.value    = Ser1$P.adj,
        threshold  = 0.05)

#  Group Letter MonoLetter
#1     1      a         a 
#2     2      b          b
#3     3      b          b
```

```{r Scar KW}
View(plotframe)

#filter for Labridae only 
Scar<-plotframe%>%filter(Family=="Scar")


hist(Scar$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Scar)
#p-value = 0.1001


ScarANOVA <- aov(density ~ dhcat, data = Scar)

# then we can easily pull them out of the output
aov_residuals <- residuals(ScarANOVA)

# let's first look to see if they look approximately normally distributed
hist(aov_residuals)

# then we can test if the residuals are approximately normally distributed with the Shapiro-Wilk Test
shapiro.test(aov_residuals) 
#nope, sad day, nan parametric it is

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Scar)

#	Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 2.0336, df = 2, p-value = 0.3618

```

```{r Pom KW}
View(plotframe)

#filter for Labridae only 
Pom<-plotframe%>%filter(Family=="Pom")


hist(Pom$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Pom)
# p-value = 1.726e-10

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Pom)

#	Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 5.1612, df = 2, p-value = 0.07573
```

```{r Mon KW}
View(plotframe)

#filter for Labridae only 
Mon<-plotframe%>%filter(Family=="Mon")


hist(Mon$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Mon)
#-value = 0.03855

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Mon)

#Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 3.6407, df = 2, p-value = 0.162
```


```{r Lut KW}
View(plotframe)

#filter for Labridae only 
Lut<-plotframe%>%filter(Family=="Lut")


hist(Lut$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Lut)
#p-value < 2.2e-16

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Lut)

#	Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 20.804, df = 2, p-value =
#3.038e-05

#now we run post hoc analysis
LutDunn=dunnTest(density ~ dhcat,
  data = Lut)

LutDunn

#  Comparison        Z      P.unadj        P.adj
#1      1 - 2 2.795159 5.187424e-03 1.037485e-02
#2      1 - 3 4.390115 1.132909e-05 3.398728e-05
#3      2 - 3 1.188514 2.346311e-01 2.346311e-01

Lut1 = LutDunn$res

cldList(comparison = Lut1$Comparison,
        p.value    = Lut1$P.adj,
        threshold  = 0.05)

#  Group Letter MonoLetter
#1     1      a         a 
#2     2      b          b
#3     3      b         b
```


```{r Lut KW}
View(plotframe)

#filter for Labridae only 
Lut<-plotframe%>%filter(Family=="Lut")


hist(Lut$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Lut)
#p-value < 2.2e-16

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Lut)

#	Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 20.804, df = 2, p-value =
#3.038e-05

#now we run post hoc analysis
LutDunn=dunnTest(density ~ dhcat,
  data = Lut)

LutDunn

#  Comparison        Z      P.unadj        P.adj
#1      1 - 2 2.795159 5.187424e-03 1.037485e-02
#2      1 - 3 4.390115 1.132909e-05 3.398728e-05
#3      2 - 3 1.188514 2.346311e-01 2.346311e-01

Lut1 = LutDunn$res

cldList(comparison = Lut1$Comparison,
        p.value    = Lut1$P.adj,
        threshold  = 0.05)

#  Group Letter MonoLetter
#1     1      a         a 
#2     2      b          b
#3     3      b         b
```


```{r Labm KW}
View(plotframe)

#filter for Labridae only 
Labm<-plotframe%>%filter(Family=="Labm")


hist(Labm$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Labm)
#p-value = 7.084e-10

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Labm)

#	Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 6.8654, df = 2, p-value = 0.0323#

#now we run post hoc analysis
LabmDunn=dunnTest(density ~ dhcat,
  data = Labm)

LabmDunn

#  Comparison         Z    P.unadj      P.adj
#1      1 - 2 1.7134507 0.08662966 0.17325931
#2      1 - 3 2.4806467 0.01311443 0.03934329
#3      2 - 3 0.5474016 0.58410288 0.58410288

Labm1 = LabmDunn$res

cldList(comparison = Labm1$Comparison,
        p.value    = Labm1$P.adj,
        threshold  = 0.05)

# Group Letter MonoLetter
#1     1      a         a 
#2     2     ab         ab
#3     3      b          b
```


```{r Car KW}
View(plotframe)

#filter for Labridae only 
Car<-plotframe%>%filter(Family=="Car")


hist(Car$density)

#bartlet test for equal vaiance

bartlett.test(density~dhcat, Car)
#-value < 2.2e-16

# too many zeros run non parametric equivalent

kruskal.test(density~dhcat, data=Car)

#Kruskal-Wallis rank sum test

#data:  density by dhcat
#Kruskal-Wallis chi-squared = 5.1013, df = 2, p-value = 0.07803
```

```{r}

sigs<-c("a","a","a","a","a","b","a","a","b","a","ab","b","a","b","b","a","a","a","a","a","a","a","a","a","a","b","b")

FNativeVsNonPlot <- ggplot(data = plotframe, aes(x = Family, y = density, fill = dhcat))+
  geom_col(position = position_dodge(0.9), width = 0.75)+
  labs(x = "Family", y = "Densities", fill = "Categories")  

FNativeVsNonPlot


```


```{r Separating Snappers}
library(readr)
library(tidyverse)
library(vegan)

AllMerged<- read_csv("AllMerged.csv")

snap<-AllMerged %>%
  group_by(Transect)%>%
  count(Common.Names)

snap<-snap%>%
  pivot_wider(names_from = Common.Names , values_from = n)

##Turning NAs to 0s
snap[is.na(snap)] <- 0

#keep only the snappers
snap<- snap %>% select("Yellowtail Snapper", "Lane Snapper", "Cubera Snapper")

#calculate density of each species on each transect
snap<-snap%>%mutate(ysdens=`Yellowtail Snapper`/100, lsdens=`Lane Snapper`/100, csdens=`Cubera Snapper`/100)

#select for density columns only
snap<-snap%>%select(Transect,ysdens, lsdens, csdens)

snap<-snap%>%
  pivot_longer(cols = ysdens:csdens,
               names_to = "Species",
               values_to = "density")

fish_den<- read_csv("Data/WeFixedItAnalyzeMe.csv")

fish_den<-fish_den %>%
  mutate(totalgrass= (fish_den$PCovHalophila+fish_den$PCovThalassia+fish_den$PCovHalodule+fish_den$PCovSyringodium)) %>%
  mutate(PerHalTotal=(fish_den$PCovHalophila/totalgrass))

##Making the NAs 0
fish_den[is.na(fish_den)] <- 0


##Make a depth column##
 fish_den <- fish_den %>%
       mutate(dcat = case_when((depth >= 15.0) ~ "deep",
                                (depth <= 15.0 & PerHalTotal == 1.0) ~ "hal",
                                (depth <= 15.0 & PerHalTotal < 1.0) ~ "native" ))

snapsbro<- merge(fish_den, snap, by= "Transect")

snapsbro<-snapsbro%>%select(Transect,)

snapsbroplot<-ggplot(data=snapsbro, aes(x=dcat, y= density, fill= Species)) +
geom_bar(stat="identity", position=position_dodge())
  #geom_errorbar( aes(x=name, ymin=value-sd, ymax=value+sd), width=0.4, colour="orange", alpha=0.9, size=1.3)


snapsbroplot

snapsbroy<-snapsbro %>% filter(Species == "ysdens")
kruskal.test(snapsbroy$density ~ snapsbroy$dcat, data = snapsbroy)

  ##Kruskal-Wallis rank sum test
  ##data:  snapsbroy$density by snapsbroy$dcat
  ##Kruskal-Wallis chi-squared = 1.0818, df = 2, p-value = 0.5822

snapsbrol<-snapsbro%>%filter(Species=="lsdens")
kruskal.test(snapsbroy$density ~ snapsbroy$dcat, data = snapsbrol)

```

