---
title: "GIS Project"
author: "Kaitlin Rommelfanger"
date: "2023-11-30"
output: html_document
---

# NMDS
This first section is the nMDS looking at fish communities at the family level, we split the seagrass into three categories, the first is all transects deeper than 15 meters, this was determined to be the depth that native seagrasses stop growing

We split the seagrass shallower than 15m into two categories 1, where all seagrass present was halophila (<15m,100% halophila), and one where there was native seagrasses present (<15m, native present)

### Three seagrass categories transects were split into
1. >15m
2. <15m, 100% Halophila
3. <15m, native present

We ran an nMDS, Permanova, and Simper to tease apart what families are most influencing differences in fish communites in these different seagrass categories
```{nMDS}
#libraries
library(tidyverse)
library(vegan)#run nMDS
library(colorBlindness)#to check plots are color blind friendly

#csv's required
pcov<-read.csv("Data/WeFixedItAnalyzeMe.csv")#seagrass info
AllMerged <- read.csv("AllMerged.csv")#fish families

#Want to know, of the seagrass present, what percentage of that seagrass is halophila, Add new column for % halophila/total seagrass cover#
pcov<-pcov %>% 
  mutate(totalgrass= (pcov$PCovHalophila+
                      pcov$PCovThalassia+
                      pcov$PCovHalodule+
                      pcov$PCovSyringodium)) %>% 
  mutate(PerHalTotal=(pcov$PCovHalophila/totalgrass))

##Nas were generated on three transects where there was no seagrass, because cannot divide by zero, Making the NAs 0
pcov[is.na(pcov)] <- 0

#rename the dataframe
depth_hal<-pcov

##Make the depth bin categories, we know that deeper than 15 meters there is no native seagrasses, and we want to separate the transects below 15 meters into two groups, ones that are 100% halophila, and ones that are not 100% halophila and contain native seagrasses ##
depth_hal<- depth_hal%>% 
       mutate(dhcat = case_when((depth >= 15.0) ~ "deep",
                                (depth <= 15.0 & PerHalTotal == 1.0) ~ "hal",
                                (depth <= 15.0 & PerHalTotal < 1.0) ~ "native" ))


##now we need to link the fish families with the seagrass information

#make family a factor
AllMerged$Family<-as.factor(AllMerged$Family)

#determine the total number of fish of each family that are on each transect to make dataframe for nMDS, which needs to be wide
dhnMDSframe<-AllMerged %>%
  group_by(Transect)%>%
  count(Family)%>%
  pivot_wider(names_from = Family , values_from = n)
 

##Turning NAs to 0s where there are no fish from that family on the transect
dhnMDSframe[is.na(dhnMDSframe)] <- 0

##Merging the two dataframes to make one tabular by transect, that includes the seagrass categories
dhnMDSframe = merge(dhnMDSframe, depth_hal, by = "Transect")


#Select for only the columns that contain the families so we can run the nMDS, want no grouping variables here
  HnMDSframe<-dhnMDSframe%>%
    select(`Aulostomidae `:Haemulidae)

#make a frame that has transect and our cat column in it to use for plotting purposes  
plotframe<-dhnMDSframe%>%
  select(`Aulostomidae `:Haemulidae, Transect,dhcat)


###Question, i need to set the seed here so we get the same nMDS for all the times we run it

#now run the nMDS 
final_NMDS <- metaMDS(HnMDSframe, k = 2, trymax = 100) 
final_NMDS
#Stress:     0.2339253 

#run the stress plot
stressplot(final_NMDS)
#non metric r squared 0.945
#linear fit r2=0.72

##pulling the species locations out of the NMDS (x and y)
fam <- as.data.frame(final_NMDS$species)


# this line is pulling the species names and linking them to thier location
fam$sp <- rownames(final_NMDS$species)


#this is pulling the points for each individual transect out of the nmds
  transect <- as.data.frame(final_NMDS$points)

#we are pulling the seagrass categories out of our plotframe that we made earlier so we can make the 95% confidence intervals for our categories depth/percent halophila
  transect$dhcat <- plotframe[,21] 

#need to keep this code because you cant pull the points in the next line unless if this ordiplot exists
ordiplot(final_NMDS, type = "none") 

#this is storing the info to make the 95% confidence intervals as an object from the ordiplot
finallord<-ordiellipse(final_NMDS, transect$dhcat, conf = 0.95, label = TRUE)

#hidden vegan function that calculates ellipses
veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
{
  theta <- (0:npoints) * 2 * pi/npoints
  Circle <- cbind(cos(theta), sin(theta))
  t(center + scale * t(Circle %*% chol(cov)))
}

#convert zone to a factor
transect$dhcat <- factor(transect$dhcat)

#create a dataframe of points that will draw the ellipses
final_df_ell <- data.frame()

for(g in levels(transect$dhcat)){
  final_df_ell <- rbind(final_df_ell, cbind(as.data.frame(with(transect[transect$dhcat==g,],
                                                   veganCovEllipse(finallord[[g]]$cov,finallord[[g]]$center,finallord[[g]]$scale)))
                                ,zone=g))
}

#plot
finalnMDSplot<-ggplot(transect) +
  geom_point(aes(x = MDS1, y = MDS2, color = as.factor(dhcat)), size = 3) + geom_path(data = final_df_ell, aes(x = NMDS1, y = NMDS2, group = zone, color=zone), linewidth = 1, linetype = 1)+
  theme_classic() +
  labs(color = "dhcat", fill = "dhcat") +
  guides(color = guide_legend(nrow = 1), fill = guide_legend(nrow = 1)) +
theme(panel.border = element_rect(fill = NA),
legend.position = "bottom")+ geom_text(data = fam, aes(x = MDS1, y = MDS2, label = sp), alpha = 0.3)+scale_fill_manual(name = "Category", labels = c("> 15m", "100% Halophila < 15m","Native Present and < 15m"),values = c("#56B4E9","#F0E442","#CC79A7"))+scale_color_manual(name = "Category", labels = c("> 15m", "100% Halophila < 15m","Native Present and < 15m"), values = c("#56B4E9","#F0E442","#CC79A7"))+annotate(geom="text", x=1.5, y=1.2, label="stress=0.234")


finalnMDSplot

#check our colors for color blind friendly
cvdPlot(finalnMDSplot)
```

```{r Permanova}

#now run post hoc to determine significant differences between different seagrasses

#rename dataframe
final_dissim_bray<- HnMDSframe

##Making the dissimilarity matrix##
final_bray <- vegdist(final_dissim_bray, method = "bray")
    final_bray

#change dhcat to factor
plotframe$dhcat<-as.factor(plotframe$dhcat)

##running the PERMANOVA# (dissimilaritymatrix~independent variables)
    dh_Strata <- adonis2(final_bray ~ plotframe$dhcat)
    dh_Strata
 
#there is a significant difference, so we run the simper to determine what is contributing to these differences 
  
###Question how to finalize the simper, some families are borderline significant, sometimes when you run it they are , and sometime they are not what is the protocol to make this
    

```

```{r Simper}
#Final simper
       dh_simper <- simper(final_dissim_bray, permutations = 999, group = (plotframe$dhcat))
  summary(dh_simper)
  

  
#deep= greater than 15m
#hal= Less than 15m and all seagrass present is Halophila
#native= Less than 15m, not all seagrass is Halophila
  

#native_deep is significant, simper shows that shallow seagrass containing native species has significantly different Lutjanidae, Serranidae, Labrisomidae (? stef and aves got significant I did not) than deep
  
#native-hal is significant simper shows that seagrass with native is significantly different Labridae, Pomacentridae, Syngnathidae, Chaetodontidae, Holocentridae than shallow 100% halophila
  
  
#deep_hal no significant families different in shallow and deep halophila

```
```

